Geometric variational problems

### 2020-11-13 NSF2020/Description/background.tex

To define HAP, consider a collection of particles suspended in a
solvent. The region $\Omega \subset \mathbb{R}^3$ models the solvent
phase (Figure \ref{fig:domain}).
%The HAP must be an energy of the shape of the region because hydrophobic
%attraction is non-additive.
The boundary of the region is the union of water-particle interfaces with
unit normal $\boldsymbol{\nu}$. Some parts of this interface are
hydrophobic while others are hydrophilic. Hydrophobic interfaces disturb
the hydrogen bond structure of water~\cite{Luzar1987, Jonsson2006,
Varilly2011}. This disturbance comes with an energetic penalty that is
proportional to interfacial area. There is an additional decay length,
due to rapid fluctuation in the hydrogen bond network, describing how
far the restructuring extends into bulk water.

@@

To establish an attraction between interfaces, the water activity is not
arbitrary, but rather is the solution of the screened Laplace
equation boundary value problem 
\begin{equation}
  \label{SL}
  -\rho^2 \Delta u + u = 0, \mbox{ } \mathbf{x} \in \Omega, \qquad
  u = f,  \mbox{ } \mathbf{x} \in \partial \Omega, \qquad 
  u(\mathbf{x}) \to 0 \mbox{ as } |\mathbf{x}| \to \infty.
\end{equation}
The boundary values $f$ define the degree of hydrophobicity of the
water-particle interface: $f=1$, describes a hydrophobic interface and
$f=0$ describes a hydrophilic interface. Thus $f$ encodes information
about the particles and the parameters $\rho$ and $\gamma$ encode
information about the quality of the solvent~\cite{Israelachvili1954,
Discher2002}. In practice, we integrate~\eqref{HAP} by parts and
use~\eqref{SL} to obtain
\begin{equation}
\label{eq:HAP_easy}
\Phi = -\gamma \int_{\partial \Omega} \rho u \nabla u \cdot \boldsymbol{\nu} \,\dif S,
\end{equation}
thereby avoiding calculating a volumetric integral.

@@

symmetric, rank-two tensor called the hydrophobic stress (eq. 2.3 in \cite{Fu2018_SIAM}):

@@

reveals the hydrophobic force and torque on each particle 

@@

relative to the center of mass $\mathbf{a}_i$.  This system is force-
and torque-free (\S\ref{subsec:specific_aim_2}).  To avoid particle

@@

overlap. The total potential,
force and torque are then

@@

As a proof of concept, our work has already tested for elastic energies
for bending, stretching, and tilt of the bilayer assembly.  The elastic
coefficients derived from the HAP simulations show strikingly positive
agreement with experimentally determined values~\cite{Fu2018_SIAM}.
Encouraged by these results and the hydrodynamic simulations of the
bilayer assembly, the PIs propose to extend the HAP model to make direct
comparison with experimental results  \S\ref{subsec:specific_aim_1}, to
establish that the HAP model has the capability and adaptivity to model
flow phenomena across length scales and time scales
\S\ref{subsec:specific_aim_2}, using efficient, high-order numerical
methods \S\ref{subsec:specific_aim_3}.

>>>>

To define HAP, consider a collection of particles suspended in a
solvent. The region $\Omega \subset \mathbb{R}^3$ models the solvent
phase (Figure \ref{fig:domain}).
%The HAP must be an energy of the shape of the region because hydrophobic
%attraction is non-additive.
The boundary of the region is the union of water-particle interfaces
with unit normal $\boldsymbol{\nu}$. Some parts of this interface are
hydrophobic while others are hydrophilic. Hydrophobic interfaces disturb
the hydrogen bond structure of water~\cite{Luzar1987, Jonsson2006,
Varilly2011} and this disturbance comes with an energetic penalty that
is proportional to interfacial area. There is an additional decay
length, due to rapid fluctuation in the hydrogen bond network,
describing how far the restructuring extends into bulk water.

@@

To establish an attraction between interfaces, the water activity is not
arbitrary, but rather is the solution of the screened Laplace
equation boundary value problem 
\begin{equation}
  \label{SL}
  -\rho^2 \Delta u + u = 0, \mbox{ } \mathbf{x} \in \Omega, \qquad
  u = f,  \mbox{ } \mathbf{x} \in \partial \Omega, \qquad 
  u(\mathbf{x}) \to 0, \mbox{ as } |\mathbf{x}| \to \infty.
\end{equation}
The boundary values $f$ define the degree of hydrophobicity of the
water-particle interface: $f=1$, describes a hydrophobic interface and
$f=0$ describes a hydrophilic interface. Thus $f$ encodes information
about the particles and the parameters $\rho$ and $\gamma$ encode
information about the quality of the solvent~\cite{Israelachvili1954,
Discher2002}. In practice, we integrate~\eqref{HAP} by parts and
use~\eqref{SL} to obtain
\begin{equation}
\label{eq:HAP_easy}
\Phi = -\gamma \int_{\partial \Omega} \rho u \nabla u \cdot \boldsymbol{\nu} \,\dif S,
\end{equation}
thereby avoiding volume integral calculations.

@@

symmetric, rank-two tensor called the hydrophobic stress (equation (2.3)
in~\cite{Fu2018_SIAM}):

@@

reveals the hydrophobic force, and torque on each particle 

@@

relative to the center of mass $\mathbf{a}_i$. This system is force- and
torque-free (\S\ref{subsec:specific_aim_2}). To avoid particle

@@

overlap. The total potential, force and torque are then

@@

As a proof of concept, our work has already tested for elastic energies
for bending, stretching, and tilt of the bilayer assembly. The elastic
coefficients derived from the HAP simulations show strikingly positive
agreement with experimentally determined values~\cite{Fu2018_SIAM}.
Encouraged by these results and the hydrodynamic simulations of the
bilayer assembly, the PIs propose to extend the HAP model to make direct
comparisons with experimental results \S\ref{subsec:specific_aim_1}, to
establish that the HAP model has the capability and adaptivity to model
flow phenomena across length scales and time scales
\S\ref{subsec:specific_aim_2}, using efficient, high-order numerical
methods \S\ref{subsec:specific_aim_3}.

### 2020-11-13 NSF2020/Description/preliminarywork.tex

\begin{wrapfigure}[13]{l}{0.30\textwidth}
\centerline{\includegraphics[width=0.30\textwidth]{figures/BIError.pdf}}
  \vspace{-8pt}
\caption{
\label{fig:bierror}  
\footnotesize The false color map shows how numerical quadrature of
  layer potentials loses accuracy near the particle boundaries.  The
  color bar is for $\log_{10}.$}
\end{wrapfigure}
\textbf{Novel reciprocal identities.} 
The force and torque formulas~\eqref{forceandtorque} require the
hydrophobic stress along the particle boundaries. Unfortunately,
standard quadrature formulas to estimate $\nabla u$ on the particle
boundaries introduces large amounts of quadrature error
(Figure~\ref{fig:bierror}). Therefore, it is useful to devise reciprocal
identities for the force and torque on body $i$ that does not
require integration along body $i$. One identity, that we have proved,
is
\begin{align}
    \label{eq:reciprocal}
{\bf F}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i} ({\bf
  x}-\mathbf{a}_i) \times [\boldsymbol{\sigma}_{ij} +
  \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, 
\end{align}
for $i=1,\ldots,N$. Here, $u_i$ is the solution of the screened Laplace
equation when only the contribution from particle $i$ is considered and
$\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i
\cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$.
Equation~\eqref{eq:reciprocal} does not require knowledge of $\nabla u$
on the particle boundary, and this is enormously beneficial for
calculating force and torque.

\subsubsection{Two-dimensional vesicles in shear flow}
We will compare the behavior of our particle-based vesicles in Stokes
flow to well-established models that enforce area incompressibility and
volume conservation in vesicle membrane
dynamics~\cite{torres-sanchez_millan_arroyo_2019,
mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}. Our
preliminary results show that the particle-based bilayers have the same
large area modulus as real lipid bilayer membranes, so we expect
realistic results from our flow simulations. Under moderate shear rates,
the vesicle elongates and takes on the tank-treading elliptical shape.
We can directly check for area compressibility conservation properties
from the particle simulation by tracking the change in area density of
the midplane as a function of shear flow rate, along with evaluating the
tangential divergence of the velocity.

>>>>

\begin{wrapfigure}[13]{l}{0.30\textwidth}
\centerline{\includegraphics[width=0.30\textwidth]{figures/BIError.pdf}}
  \vspace{-8pt}
\caption{
\label{fig:bierror}  
\footnotesize The false color map shows how numerical quadrature of
  layer potentials loses accuracy near the particle boundaries.  The
  color bar is for $\log_{10}.$}
\end{wrapfigure}
\textbf{Novel reciprocal identities.} 
The force and torque formulas~\eqref{forceandtorque} require the
hydrophobic stress along the particle boundaries. Unfortunately,
standard quadrature formulas to estimate $\nabla u$ on the particle
boundaries introduces large amounts of quadrature error
(Figure~\ref{fig:bierror}). Therefore, it is useful to devise reciprocal
identities for the force and torque on body $i$ that does not
require integration along body $i$. One identity, that we have proved,
is
\begin{align}
    \label{eq:reciprocal}
{\bf F}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i}
[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i} ({\bf
  x}-\mathbf{a}_i) \times [\boldsymbol{\sigma}_{ij} +
  \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, 
\end{align}
for $i=1,\ldots,N$. Here, $u_i$ is the solution of the screened Laplace
equation when only the contribution from particle $i$ is considered and
$\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i
\cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$.
Because of the double layer potential jump relations, \eqref{eq:reciprocal} 
does not require knowledge of $\nabla u$ on the particle boundary, and this
is enormously beneficial for calculating force and torque.

\subsubsection{Two-dimensional vesicles in shear flow}
We have simulated two-dimensional vesicles in the background shear flow
$\mathbf{u}_{\infty} = \dot{\gamma} x_2 \mathbf{i}_1$. The numerical solver
is written in the Matlab language runs on a desktop without parallelization. 
The initially circular vesicle bilayers rotate and elongate with the shear flow, 
and approach a steadily tank-treading ellipse.

There are a number of well-established mathematical models of vesicles
in shear flow \cite{torres-sanchez_millan_arroyo_2019, mahapatra_saintillan_rangamani_2020, 
Steigmann99, C6SM02452A}. We are trying to assess which constituative
assumption the particle-based bilayer satisfies. 
Our preliminary results show that the particle-based bilayers have the same
large area modulus as real lipid bilayer membranes, so we expect
realistic results from our flow simulations. Under moderate shear rates,
the vesicle elongates and takes on the tank-treading elliptical shape.
We can directly check for area compressibility conservation properties
from the particle simulation by tracking the change in area density of
the midplane as a function of shear flow rate, along with evaluating the
tangential divergence of the velocity.
